Synthesis of Linear Functional Programs

### 2021-05-09 article/article.tex
tweaks

\mypara{Propositions as Types}
%
It turns out that the inference rules of the simply-typed
$\lambda$-calculus are closely related to those of a system of natural
deduction for intuitionistic logic~\cite{prawitznd65}. This
relationship, known as the Curry-Howard correspondence~\cite{},
identifies that the propositions of intuitionistic logic can be read
as types for the simply-typed $\lambda$-calculus (``propositions as
types''), their proofs are exactly the program with the given type
(``proofs as programs''), and checking a proof is type checking a
program (``proof checking as type checking'')~\cite{}.

@@

  The elimination form for the exponential, written $\vdash \llet{ !u
    = M}{ N}$, warrants the use of the resource $A$ an unbounded
  number of times in $N$ via the variable $u$:  
\[
    \infer*[right=($\bang E$)]
    {\Gamma ; \Delta_1 \vdash M : \bang A \and \Gamma, u{:}A ; \Delta_2 \vdash N : C}
    {\Gamma ; \Delta_1, \Delta_2 \vdash \llet{ !u = M}{ N} : C}
\]

@@

\mypara{Synthesis from (polymorphic) refinement types}~\cite{DBLP:conf/pldi/PolikarpovaKS16}
This paper also attempts to synthetize recursive functional programs, in a more ``advanced'' setting. The specification
is a combination of two rich types, polymorfic and refinement types (which correspond to a first-order logic through the Curry-Howard isomorphism) -- this is an interesting
combo that offers ``expressive power and decidability, which enables automatic verification--and hence synthesis--of nontrivial programs''.
Their approach to the refinement types is a new algorithm that supports decomposition of the refinement specification,
and they separation between the language of specification and programs.

\mypara{Synthesis for graded types} (no recursion   -- in that work, no
  obvious connection of grading with
  concurrency)~\cite{DBLP:conf/lopstr/HughesO20}

>>>>

\mypara{Propositions as Types}
%
It turns out that the inference rules of the simply-typed
$\lambda$-calculus are closely related to those of a system of natural
deduction for intuitionistic logic~\cite{prawitznd65}. This
relationship, known as the Curry-Howard
correspondence~(see~\cite{DBLP:journals/cacm/Wadler15} for a
historical survey),
identifies that the propositions of intuitionistic logic can be read
as types for the simply-typed $\lambda$-calculus (``propositions as
types''), their proofs are exactly the program with the given type
(``proofs as programs''), and checking a proof is type checking a
program (``proof checking as type checking'').

@@

  The elimination form for the exponential, written $\llet{ !u
    = M}{ N}$, warrants the use of resource $A$ an unbounded
  number of times in $N$ via the variable $u$:  
\[
    \infer*[right=($\bang E$)]
    {\Gamma ; \Delta_1 \vdash M : \bang A \and \Gamma, u{:}A ; \Delta_2 \vdash N : C}
    {\Gamma ; \Delta_1, \Delta_2 \vdash \llet{ !u = M}{ N} : C}
\]

@@

\mypara{Program Synthesis from Polymorfic Refinement Types} The
work~\cite{DBLP:conf/pldi/PolikarpovaKS16} also studies synthesis
recursive functional programs but in a more ``advanced'' setting. The
specification is a combination of two rich forms of types: polymorphic and
refinement types (which correspond to a first-order logic through the
Curry-Howard isomorphism) -- this is an interesting combination that offers
``expressive power and decidability, which enables automatic
verification--and hence synthesis--of nontrivial programs''.  Their
approach to refinement types consists a new algorithm that supports
decomposition of the refinement specification, allowing for separation
between the language of specification and programs and making the
approach amenable to compositional synthesis techniques.

\mypara{Resourceful Program Synthesis from Graded Linear Types} The
work~\cite{DBLP:conf/lopstr/HughesO20} synthesizes programs using an
approach very similar to our own.  The work employs so-called graded
modal types, which is an refinement of pure linear types -- a more
\emph{fine-grained} version, since it allows for quantitative
specification of resource usage, in contrast to ours either
\emph{linear} or \emph{unrestricted} (via the linear logic
exponential) use of assumptions.  Their resource management is more
complex, and so they provide solutions which adapt Hodas and Miller's
resource management approach~\cite{DBLP:journals/tcs/CervesatoHP00,DBLP:journals/tcs/LiangM09} -- the model we'll be using.

They also use focusing as a solution to trim down search space and
to ensure that synthesis only produces well-typed programs, and
our solution will use the same literature as a base.  However, since
their underlying logic is \emph{modal} rather than purely
\emph{linear}, it lacks a clear correspondence with concurrent
session-typed programs~\cite{}, which is a crucial avenue of future
work. Moreover, their use grading effectively requires constraint solving to
be integrated with the synthesis procedure, which can limit the
effectiveness of the overall approach as one scales to more
sophisticated settings (e.g.~refinement types).

% (no recursion   -- in that work, no
%   obvious connection of grading with
%   concurrency)
